Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-09 manuscript.Rmd
first half of distal is done

Figure \@ref(fig:part2overall) and Table \@ref(tab:part2overalltable) summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded prevalence was *r* = `r split_agree_cor`. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of \@ref(fig:part2overall) indicated two extreme outliers, marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Infant was in a baby walker for most of part 2. Moved more like a baby held by a CG than a baby walking upright.

For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.

>>>>

Figure \@ref(fig:part2overall) and Table \@ref(tab:pt2overalltable) summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded prevalence was *r* = `r split_agree_cor`. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of \@ref(fig:part2overall) indicated two extreme outliers, marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs. sitting; upright vs. held) and their accuracy for other classes was strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \@ref(tab:pt2overalltable) excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). With the exception of held, within-class agreement was strong for the other four body positions.

### 2023-03-09 manuscript.tex
first half of distal is done

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}). As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019). For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

@@

@@

What can we learn about the two outliers? What do they tell us about how this will work in the future.

One outlier (106-4) had significant confusion between sitting and supine. Both their proximal and distal performance was bad, but only for sitting and supine. Watching the videos showed that this participant spent a long period of time in a seating device that was reclined between supine and sitting. Also spent long time in mother's arms in a similar supine/sitting position that was hard to classify.

The other outlier (107-3) had good proximal accuracy but poor distal accuracy. Confusion between upright and suspended. Infant was in a baby walker for most of part 2. Moved more like a baby held by a CG than a baby walking upright.

For both, other class performance was good, suggesting that it's about the particular events, not about the quality of the algorithm in applying to the participant overall.

>>>>

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \ref{fig:exemplar-timeline}). A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to \textasciitilde{} 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work (Franchak et al., 2021; Yao et al., 2019). For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

@@

Figure \ref{fig:part2overall} and Table \ref{tab:part2overalltable} summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded prevalence was \emph{r} = 0.91. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of \ref{fig:part2overall} indicated two extreme outliers, marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions. The ``gray square'' outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''. Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held) and their accuracy for other classes was strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \ref{tab:part2overalltable} excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96). With the exception of held, within-class agreement was strong for the other four body positions.

@@
